An Ant Cuticle Texture Classification Algorithm for Ecological Anaylsis

### 2021-09-03 paper/paper.tex
auto generate antweb figures
add subimage dataset

Texture is an important feature in many applications, such as image processing,
pattern recognition, and computer vision. Analysis of textures can be broken
into three main categories: texture classification, texture segmentation, and
texture synthesis. The process of classifying a texture into a set of categories
and relies on three different approaches: feature-based, model-based, or
structural \cite{reed_review_1993}. In this paper, we focus on a
\textit{model-based approach} which attempt to extract parameters to reveal
common patterns and use those parameters to automatically distinguish between
different textures \cite{maillard_texture_2003}.

@@

@@

>>>>

Texture is an important feature in many applications, such as image processing,
pattern recognition, and computer vision. Analysis of textures can be broken
into three main categories: texture classification, texture segmentation, and
texture synthesis \cite{reed_review_1993}. The process of classifying a texture
into a set of categories and relies on three different approaches. In this
paper, we focus on a \textit{model-based approach} which attempts to extract
parameters to reveal common patterns and use those parameters to automatically
distinguish between different textures \cite{maillard_texture_2003}.

@@

\subsection{Subimage Dataset}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{assets/images/segmented_1.png}
  \caption{Hand segmented ant head image to separate head cuticle from other
    sections of the image.
    \href{https://www.antweb.org/bigPicture.do?name=casent0217419&shot=h&number=1}{CASENT0217419}
    by \href{https://www.antweb.org/artist.do?id=92}{Estella Ortega}, from
    \href{https://www.antweb.org}{AntWeb}}
  \label{fig:segmented_1}
\end{figure}

When using the Sculpture Identification Protocol, the class of the overall image
is based only on the texture of the head, not the eyes, antenna, or any other
exposed cuticle. We use image segmentation to analyze the texture by splitting
the ant head image into a set of regions: \textit{cuticle} and
\textit{background}. A benefit of this approach allows us to pull patches or
subimages of various sizes from the region to increase the number of samples
available for training and at the same time, reduce the number of features to
simplify the classification algorithm. With a given subimage size to extract
from the segmented image, we label each subimage extracted based on the class of
the input image. Overall, there are 73 images that were hand segmented in this
manner and the class distribution for various sizes of subimages are shown in
Table \ref{table:classdist_rs_sub}.

\begin{table}[h]
  \centering
  \caption{Subimage Dataset Class Distribution}
  \label{table:classdist_rs_sub}
  \include{tables/rs_samples_sub.tex}
\end{table}

@@

In the first experiment, we investigate the ability of a basic CNN to classify
cuticle texture into the two main categories: \textit{rough} and
\textit{smooth}. The goal of this experiment is to determine if more intricate
data preprocessing is necessary to classify the texture of the ant. The dataset
used is shown in Table \ref{table:classdist_rs_full}, and the results are shown
in (TODO).
